Probabilistic Forecasting of the Disturbance Storm Time Index: An Autoregressive Gaussian Process approach

### 2017-03-01 agutmpl.tex
Updates from ShareLaTeX

\usepackage{amsmath}
\usepackage{txfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{url}
\usepackage{lineno}
\linenumbers*[1]
\usepackage{rotating}
\usepackage{adjustbox}
\graphicspath{{figures/}}
%  To add line numbers to lines with equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables
%
%
% DO NOT USE \psfrag or \subfigure commands.
%
%  Figures and tables should be placed AT THE END OF THE ARTICLE,
%  after the references.
%
%  Uncomment the following command to include .eps files
%  (comment out this line for draft format):
%  \usepackage[dvips]{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
%  \setkeys{Gin}{draft=false}
%
% Substitute one of the following for [dvips] above
% if you are using a different driver program and want to
% proof your illustrations on your machine:
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
% See how to enter figures and tables at the end of the article, after
% references.
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

@@

%Talk about need for probabilistic forecasts.
Although much research has been done on prediction of the $Dst$ index, must less has been done on probabilistic forecasting of $Dst$. One such work described in \citet{McPherron:2013} involves identification of high speed solar wind streams using the W-S-A model (see \citet{WSAModel}), using predictions of high speed streams to construct ensembles of $Dst$ trajectories which yield the quartiles of $Dst$ time series.

In this work we propose an technique for probabilistic forecating of $Dst$, which yields a predictive distribution as a closed form expression. Our models take as input past values of $Dst$, solar wind speed and the z-component of the \emph{Interplanetary Magnetic Field} (IMF) and output a Gaussian distribution with a specific mean and variance as the OSA prediction of the $Dst$.

@@

The quantities $l$ in the RBF, and $b$ and $d$ in the polynomial kernel are known as \emph{hyper-parameters}. Hyper-parameters give flexibility to particular kernel structure, for example $d = 1, 2, 3, \cdots$ in the polynomial kernel represents linear, quadratic, cubic and higher order polynomials respectively. The process of assigning values to the \emph{hyper-parameters} is crucial in the model building process and is known as \emph{model selection}.

@@

The computational complexity of calculation of the predictive distribution is $O(N^3)$, as discussed in section \ref{sec:inference}. This can limit the size of the covariance matrix constructed from the training data. Note that this computation overhead is paid for every unique assignment to the model hyper-parameters. However, our chosen training set has a size of 243 which is still very much below the computational limits of the method and in our case. Solving equation \ref{eq:posteriormean} on a laptop computer takes less than a second for the training set considered in our analysis.

@@

In order to find appropriate values of the hyper-parameters of the chosen kernel $K_osa(.,.)$, we apply \emph{grid search}, \emph{coupled simulated annealing} and \emph{maximum likelihood} methods. We fix the parameters $d$ and $\sigma^2$ of $K_st$ and model noise to values $0.01$ and $0.2$ respectively, the remaining parameters $w$ and $b$ are kept free to be calculated by model selection. Table \ref{table:modelselection} summarizes the settings used to run each model selection procedure.

@@

For measuring performance of model instances on the validation set storm events, the \emph{mean absolute error} (MAE), \emph{root mean square error} (RMSE) and \emph{coefficient of correlation} (CC) are calculated.

@@

Figures \ref{fig:CompareMaeARX} and \ref{fig:CompareCCARX} break down the results for GP-ARX by the model selection routine used. Apart from the general trend observed in \ref{fig:CompareMAE} and \ref{fig:CompareCC}, we also observe that \emph{grid search} and \emph{coupled simulated annealing} give superior performance as compared to gradient based \emph{maximum likelihood}.

>>>>

\usepackage{amsmath}
\usepackage{txfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{url}
\usepackage{lineno}
\linenumbers*[1]
\usepackage{rotating}
\usepackage{adjustbox}
\graphicspath{{figures/}}
%  To add line numbers to lines with equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables
%
%
% DO NOT USE \psfrag or \subfigure commands.
%
%  Figures and tables should be placed AT THE END OF THE ARTICLE,
%  after the references.
%
%  Uncomment the following command to include .eps files
%  (comment out this line for draft format):
%  \usepackage[dvips]{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
  \setkeys{Gin}{draft=false}
%
% Substitute one of the following for [dvips] above
% if you are using a different driver program and want to
% proof your illustrations on your machine:
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
% See how to enter figures and tables at the end of the article, after
% references.
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

@@

%Talk about need for probabilistic forecasts.
Although much research has been done on prediction of the $Dst$ index, much less has been done on probabilistic forecasting of $Dst$. One such work described in \citet{McPherron:2013} involves identification of high speed solar wind streams using the WSA model (see \citet{WSAModel}), using predictions of high speed streams to construct ensembles of $Dst$ trajectories which yield the quartiles of $Dst$ time series.

In this work we propose a technique for probabilistic forecasting of $Dst$, which yields a predictive distribution as a closed form expression. Our models take as input past values of $Dst$, solar wind speed and the \textit{z} component of the \emph{Interplanetary Magnetic Field} (IMF) and output a Gaussian distribution with a specific mean and variance as the OSA prediction of the $Dst$.

@@

The quantities $l$ in the RBF, and $b$ and $d$ in the polynomial kernel are known as \emph{hyper-parameters}. Hyper-parameters give flexibility to a particular kernel structure, for example $d = 1, 2, 3, \cdots$ in the polynomial kernel represents linear, quadratic, cubic and higher order polynomials respectively. The process of assigning values to the \emph{hyper-parameters} is crucial in the model building process and is known as \emph{model selection}.

@@

The computational complexity of calculation of the predictive distribution is $O(N^3)$, as discussed in section \ref{sec:inference}. This can limit the size of the covariance matrix constructed from the training data. Note that this computation overhead is paid for every unique assignment to the model hyper-parameters. However, our chosen training set has a size of 243 which is still very much below the computational limits of the method and in our case solving equation \ref{eq:posteriormean} on a laptop computer takes less than a second for the training set considered in our analysis.

@@

In order to find appropriate values of the hyper-parameters of the chosen kernel $K_{osa}$, we apply \emph{grid search}, \emph{coupled simulated annealing} and \emph{maximum likelihood} methods. We fix the parameters $d$ and $\sigma^2$ of $K_{st}$ and model noise to values $0.01$ and $0.2$ respectively, the remaining parameters $w$ and $b$ are kept free to be calculated by model selection. Table \ref{table:modelselection} summarizes the settings used to run each model selection procedure.

@@

For measuring performance of model instances on the validation set storm events, the following metrics are calculated.

\begin{enumerate}
    \item The mean absolute error.
    \begin{equation}
        MAE = \sum_{t=1}^{n} \left |(Dst(t) - \hat{D}st(t)) \right | / n
    \end{equation}
    \item The root mean square error.
    \begin{equation}
        RMSE = \sqrt{\sum_{t=1}^{n} (Dst(t) - \hat{D}st(t))^2 / n}
    \end{equation}
    \item Correlation coefficient between the predicted and actual value of $Dst$.
    \begin{equation}
        CC = Cov(Dst, \hat{D}st)/\sqrt{Var(Dst) Var(\hat{D}st)}
    \end{equation}
\end{enumerate}

@@

Figures \ref{fig:CompareMaeARX} and \ref{fig:CompareCCARX} break down the results for GP-ARX by the model selection routine used. Apart from the general trend observed in \ref{fig:CompareMae} and \ref{fig:CompareCC}, we also observe that \emph{grid search} and \emph{coupled simulated annealing} give superior performance as compared to gradient based \emph{maximum likelihood}.
